---
title: "Recognized and Assumed Data Formats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recognized and Assumed Data Formats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(terradactyl)
```

# Overview
Many of the functions in terradactyl make assumptions about incoming data tabular formats, specifically the names of variables containing expected values and their data types/classes. In some cases, there are arguments for providing the name of a variable for a particular value type, e.g., `species_join()` has arguments for identifying the variables that contain species attributes like growth habits. Other functions like the `gather_*()` family have exclusively hard-coded variable names based on the incoming data types.

Some variables are common across multiple or even all data types. Most critically, this includes PrimaryKey which is expected in nearly every data input. This document describes the expected and required variables for all tabular data and which data sets they are part of.

To see the expectations for specific variables, reference the [Variables] section.

To see the expectations for a table, reference the [Data Types] section.

# Variables
The following are the variables that various functions expect, their descriptions, the allowed values, and which data types they're required for. To see these organized by data type, see the [Data Types] section.

## PrimaryKey

### Value type
Character string

### Present in
[tblLPIHeader]

### Description
PrimaryKey is intended to function as the unique identifier for a sampling event at a location, i.e., a field crew visiting a plot and collecting data with one or more methods. Every record across all data tables associated with a sampling event should share the same PrimaryKey value. Because PrimaryKey is used to group data as a sampling event, a value should be shared between data collected across multiple visits to a location *if* that is still considered one sampling event, e.g. collecting different data types across multiple days. Consider timing aspects such as phenological windows or interseasonal variability when determining what constitutes a sampling event; data collection separated by months or years should not share a PrimaryKey because combining and comparing those data would be ecologically inappropriate.

In most tables, PrimaryKey values are not unique per-record. Exceptions include header tables (e.g. [tblLPIHeader]) in which there should only be one record per sampling event at a location.

## LineKey

### Value type
Character string

### Present in
[tblLPIHeader]

### Description
LineKey functions as the unique identifier for a transect *for a [PrimaryKey]*. The values in LineKey is not expected to be a unique identifier for a transect within any table because it is intended to be used in conjunction with [PrimaryKey]. Any uniqueness will be incidental.

All records associated with the same transect—that is, collected along, from, or on that transect—share the same LineKey value. This is expected to hold true across tables. For example, canopy gap data and LPI data collected on the same transect for the same PrimaryKey would be expected to share the same LineKey value.

## LineLengthAmount

### Value type
Numeric

### Present in
[tblGapHeader]

### Description
The length of the transect for the associated [PrimaryKey] and [LineKey] combination.

The units for this variable are not standardized. In cases like [tblGapHeader] the [Measure] variable is used to indicate whether this is imperial or metric. Typically, this value is in meters but may also be recorded in centimeters or even feet.

## Measure

### Value type
Numeric

### Present in
[tblGapHeader]

### Description
This variable indicates the family of units used for recording data for a combination of method, [PrimaryKey], and [LineKey]. When the value of Measure is 1 the units are metric (centimeters or meters). When the value is 2 the units are imperial (inches or feet). The particular units used for a measurement are specific to the method and implementation.

## GapStart
Mark on transect in cm or in where a gap started. Units indicated by associated [Measure] value

## GapEnd
Mark on transect in cm or in where a gap started. Units indicated by associated [Measure] value

## Gap
Length of gap in cm or in. Units indicated by associated [Measure] value.

## RecType
"C", "B", or "P" corresponding to if the gap represents canopy gap, basal gap, or perennial-only canopy gap, respectively.


# Data Types
Data types are cool

## Cover and heights
### Tables
#### tblLPIHeader
#### tblLPIDetail
#### lpi_tall

## Canopy gaps

### Tables

#### tblGapHeader

#### tblGapDetail

#### gap_tall
This is the tall/long-format gap data used by functions like `gap_cover()` and which can be produced by `gather_gap()` from [tblGapHeader] and [tblGapDetail].

##### Variables
[PrimaryKey], [LineKey], [LineLengthAmount], [Measure], [GapStart], [GapEnd], [Gap], [RecType]

## Species inventory
### Tables
#### tblSpeciesInvHeader
#### tblSpeciesInvDetail
#### species_tall

## Soil stability
### Tables
#### tblSoilStabHeader
#### tblSoilStabDetail
#### gap_tall

## Interpreting Indicators of Rangeland Health (IIRH) {IIRH}
